---
title: "Palmer Penguins R Analysis"
author: "Jayke Sudana"
date: "1/20/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### The goal of this analysis is to practice basic R functions and data exploration on the Palmer Penguins data set.  
The learning comes from the [Google Data Analytics Certification offered on Coursera.org](https://www.coursera.org/professional-certificates/google-data-analytics).

### Setting up my environment
Notes: setting up my R environment by loading the 'tidyverse', 'palmerpenguins', 'here', 'skimr', and 'janitor' packages
```{r}
install.packages("palmerpenguins")
install.packages("tidyverse")
install.packages("skimr")
install.packages("here")
install.packages("janitor")
install.packages("ggplot2")

library(tidyverse)
library(palmerpenguins)
library(skimr)
library(here)
library(janitor)
```

### Viewing the penguins data set
```{r}
head(penguins)
```
#### Familiarizing ourselves with the penguins data set
```{r}
str(penguins)
```
#### More information on the data set
```{r}
colnames(penguins)
```

### Ordering penguins data set in descending order by bill length
```{r}
penguins %>% 
  arrange(-bill_length_mm)
penguins_2 <- penguins %>% arrange(-bill_length_mm)
view(penguins_2)
```

### Creating summary statistics of penguins data set by species and island.
```{r}
penguins %>% 
  group_by(species, island) %>% 
  drop_na() %>% 
  summarize(max_bl = max(bill_length_mm), mean_bl = mean(bill_length_mm))
```

### Filtering the data set for 'Adelie' species
```{r}
penguins %>% 
  filter(species == "Adelie")
```

### Using mutate to add columns with changed values in penguins data set
```{r}
penguins %>% 
  mutate(body_mass_kg=body_mass_g/1000, flipper_length_m=flipper_length_mm/1000)
```

### Using ggplot2 to visualize the relationship between penguin flipper length and body mass via a scatterplot
#### Setting up my ggplot2 environment
```{r}
library(ggplot2)

ggplot(data = penguins)+geom_point(mapping=aes(x = flipper_length_mm, y = body_mass_g))
```

### Adding a trend line in our scatterplot of body mass vs. flipper length
```{r}
ggplot(data=penguins)+
  geom_smooth(mapping=aes(x=flipper_length_mm, y=body_mass_g)) +
  geom_point(mapping=aes(x=flipper_length_mm, y=body_mass_g))
```

### Displaying penguin body mass vs flipper length with a different linetype for each species.
```{r}
ggplot(data=penguins)+
  geom_smooth(mapping=aes(x=flipper_length_mm,
                          y=body_mass_g,
                          linetype=species))
```

### Using facet_wrap() to display sub plots of the penguin data set based on species.
```{r}
ggplot(data=penguins)+
  geom_point(mapping=aes(x=flipper_length_mm,y=body_mass_g,color=species))+
  facet_wrap(~species)
```

### Using facet_grid() to add an additional variable to our subplot, adding a vertical dimension to our subplots.
```{r}
ggplot(data=penguins)+
  geom_point(mapping=aes(x=flipper_length_mm,
                         y=body_mass_g,
                         color=species))+
  facet_grid(sex~species)
```

### Visualizing the penguin data sets with annotations, titles, and subtitles.
```{r}
ggplot(data=penguins)+
  geom_point(mapping=aes(x=flipper_length_mm,
                         y=body_mass_g,
                         color=species))+
  labs(title="Palmer Penguins: Body Mass vs. Flipper Length",
       subtitle="Sample of Three Penguin Species",
       caption="Data Collected by Dr. Kristen Gorman")+
  annotate("text", x=220,y=3500, label="The Gentoos are the largest")

```

### Saving the data visualizations we created
```{r}
ggsave("Palmer Penguins Body Mass vs Flipper Length.png")
```


